---
title: "Lab 1: NYT API"
author: "Sofia Ingersoll"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite) #convert results from API queries into R-friendly formats 
library(tidyverse) #tidy
library(tidytext) #text data management and analysis
library(ggplot2) #plot word frequencies and publication dates

#assign API key.  When you create a NYT Dev account, you will be given a key
API_KEY # stored in key.Rmd
```

Today we will be grabbing some data from the New York Times database via their API, then running some basic string manipulations, trying out the tidytext format, and creating some basic plots.

<https://developer.nytimes.com/>

### Connect to the New York Times API and send a query

We have to decide which New York Times articles we are interested in examining. For this exercise, I chose articles about Deb Haaland, the current US Secretary of the Interior. As a member of the Laguna Pueblo Tribe, Haaland is the first Native American to serve as a Cabinet secretary. Very cool!

We'll send a query to the NY Times API using a URL that contains information about the articles we'd like to access.

fromJSON() is a wrapper function that handles our request and the API response. We'll use it to create an object,t, with the results of our query. The flatten = T argument converts from the nested JSON format to an R-friendlier form.

```{r api, eval = FALSE}

#create the query url
url <- paste("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=haaland&api-key=",API_KEY, sep ="")

#send the request, receive the response, and flatten
t <- fromJSON(url, flatten = T)
```

```{r api, eval = FALSE}
#what type of object is it?
#str(t)

#convert to a data frame 
t <- as.data.frame(t)

# how big is it?
#str(t)
# what fields are we working with?

```

The name format, "response.xxx.xxx...", is a legacy of the JSON nested hierarchy.

Let's look at a piece of text. Our data object has a variable called "response.docs.snippet" that contains a short excerpt, or "snippet" from the article. Let's grab a snippet and try out some basic string manipulations from {stringr}.

```{r basic_stringr, eval=FALSE}
t$response.docs.snippet[9]

# assign a snippet to x to use as fodder for stringr functions.  

x <- "Her nomination as secretary of the interior is historic, but as the first Native cabinet member, she would have to strike a delicate balance." 

# convert to lowercase.  
tolower(x)

# split into substrings
str_split(x, ',')

# swap strings
str_replace(x, "historic", "without precedent")

#how do we replace all of them?
str_replace_all(x, " ", "_")

# detect a string: returns boolean
str_detect(x,"t")

# locate it
str_locate(x, "t")

str_locate_all(x, 'as')
```

### OK, it's working but we want more data. Let's set some parameters for a bigger query.

```{r}
term1 <- "Deb" 
# Need to use & to string  together separate terms
term2 <- "&Haaland" 

begin_date <- "20210120"
end_date <- "20230401"

#construct the query url using API operators
baseurl <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",term1,term2,"&begin_date=",begin_date,"&end_date=",end_date,"&facet_filter=true&api-key=",API_KEY, sep="")

#examine our query url
baseurl
```

The initial query returned one page of ten articles, but also gave us count of total number of hits on our query. We can use that to size a for() loop to automate requests.

```{r, eval=FALSE}
#run initial query
initialQuery <- fromJSON(baseurl)

# limit the number of articles returned per request
# use this for the lab to find how many avaialbe hits are on my query so far to make sure how far to go thru the loop to collect everything
#maxPages <- round((initialQuery$response$meta$hits[1] / 10)-1) 

# for our demo, only want things with 10 hits for simplicity
maxPages <- 10
#initiate a list to hold results of our for loop
pages <- list()
# loop to automate requests
# API urls are zero indexed, that's why we're starting at 0 here to return a total of 11 articles
# each item in the list is a column containing lists
for(i in 0:maxPages){
  nytSearch <- fromJSON((paste0(baseurl, "&page=", i), flatten = TRUE) %>% data.frame() 
  # updating us on scrapping process
  message("Retrieving page ", i)
  # storing results in the list
  pages[[i+1]] <- nytSearch 
  
  # API is limited by query request size
  Sys.sleep(20)
}
```

We converted each returned JSON object into a data frame.

```{r, bind_rows}
# bind the pages and create a tibble from nytDat
nyt_df  <- bind_rows(pages)

nyt_df
```

Let's start exploring our data.  What types of content did we turn up?
```{r article-type}
colors = c("pink", "orange", "plum")

nyt_df %>% 
  # we want type_of_material so we need to navigate thru response.docs
  group_by(response.docs.type_of_material) %>%
  # This creates a new data frame with the count of records for each type_of_material.
  summarize(count=n()) %>% 
  # add percent of total column
  mutate(percent = (count / sum(count))*100) %>% 
  
  ggplot() +
  
  geom_bar(aes(y=percent,
               x=response.docs.type_of_material,
               fill=response.docs.type_of_material),
           stat = "identity") + 
  coord_flip() #+ 
  #scale_color_manual(values = colors )
```

```{r date-plot}
nytDat <- nyt_df

# how does the sentiment of the news change over the tine?
nytDat %>%
  # globally substitute all occurances
  # pull out first part of the date and we're not interested in the time portion
  mutate(pubDay=gsub("T.*","",response.docs.pub_date)) %>% 
  
  
  group_by(pubDay) %>%
  
  summarise(count=n()) %>%
  
 # filter(count >= 2) %>%
  
  ggplot() +
  
  geom_bar(aes(x=reorder(pubDay, count), y=count), stat="identity") +
  #bring date so bars go longwise
  coord_flip() 
```

The New York Times doesn't make full text of the articles available through the API. But we can use the first paragraph of each article.

```{r plot_frequencies}
#find first paragraph field
names(nytDat)

#6 is lead_paragraph
# this is response.docs.lead_paragraph
paragraph <- head(nytDat[,1:6])


# we wanna exclude things
# token is a fundamental unit of analysis for text data
tokenized <- nytDat %>% 
  # area of nyt focusing on certain topics
  filter(response.docs.news_desk != c("Sports", "Games")) %>% 
  # reduces text down to token
  # creates data structure were each row responds to a single token
  # word is new column we want, paragraph is the source we're unnesting from
  unnest_tokens(word, response.docs.lead_paragraph)

#use tidytext::unnest_tokens to put in tidy form.  
tokenized[,"word"]
```

Alright, let's starting analyzing our data.  What the most frequent words in the articles we have?
```{r word_frequencies}
tokenized %>%
  count(word, sort = TRUE) %>% 
  filter(n > 100) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(x = n, 
             y = word )) +
  geom_col() +
  labs(y = NULL)
  
```

Uh oh, who knows what we need to do here?

```{r stop-words}
#load stop words
data("stop_words")
stop_words

#stop word anti_join
tokenized <- tokenized %>% 
  # remove stop words
  anti_join(stop_words)


#now let's try that plot again
tokenized %>%
  count(word, sort = TRUE) %>% 
  filter(n > 20) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(x = n, 
             y = word )) +
  geom_col() +
  labs(y = NULL)
```

OK, but look at the most common words. Does anything stand out?

Combining contractions with singular words
"government's + government" 
```{r}
length(tokenized)
length(clean_tokens)
```

```{r cleaning, eval=FALSE}

#inspect the list of tokens (words)
tokenized$word

# remove all of these numeric characters from this column
# these year tokens are typically now stored as empty strings
clean_tokens <- str_remove_all(tokenized$word, "[:digits:]")

# goodbye apostrophe s
clean_tokens <- gsub("'s", "", clean_tokens)

# adding clean tokens back as a column with some empty strings 
tokenized$clean <- clean_tokens

# subsetting data for clean text
tib <- subset(tokenized, clean != "")


tokenized %>%
  count(clean, sort = TRUE) %>%
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)

#try again

```

## Assignment (Due Tuesday 4/9 11:59pm)
Reminder: Please suppress all long and extraneous output from your submissions (ex:  lists of tokens).

1.  Create a free New York Times account (<https://developer.nytimes.com/get-started>)

2.  Pick an interesting environmental key word(s) and use the {jsonlite} package to query the API. Pick something high profile enough and over a large enough time frame that your query yields enough articles for an interesting examination.

3.  Recreate the publications per day and word frequency plots using the first paragraph field.  This time filter on the response.docs.news_desk variable to winnow out irrelevant results.

-   Make some (at least 3) transformations to the corpus including: add context-specific stopword(s), stem a key term and its variants, remove numbers)

4.  Recreate the publications per day and word frequency plots using the headlines variable (response.docs.headline.main). Compare the distributions of word frequencies between the first paragraph and headlines. Do you see any difference?
